Java Based Memory Latency Simulator

### 2020-11-21 finalPaper/paper.tex
Updated

Seek time; which is the time it takes for the Read-Write to move up and down the platter.
The Seek time itself is comprised of a series of phase times. The first being the acceleration phase,
the read write arm starts to move, then comes the coasting phase when the arm is moving at full speed. This
is followed by the deceleration phase where the read write arm slows down. Then finally comes the
settling phase as the read write head stops right at the data it needs.
The average seek time of hardrives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
(source: pages.cs.wisc.edu - Operating Systems: Three Easy peices peices - Remzi Arpaci, Andrea Arpaci).
(source: \cite }
The last part of total I/O latency for hard drive is the transfer time, which is the time it takes the
data to be transfer from or to the Read-Write head. Compared to Rotational Latency and seek time, transfer
time is insignificant, given that the mechanical movements take more time than data movement which is more of
movement of energy packets.

@@

In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the 'Operating Systems: Three
easy peices', two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
data being dealt with was 100 MB. However, another controller in the comparison was whether the data read or writing
was sequential or random. Sequential meant that large chunks of data are just being read and written which are in
sequence, whereas the Random transfer just meant Random data was selected from a read issue some where on the disk,
which is how the data for applications work. Mind it that the data we represented in the simulator was also sequential
and not random, as the hardrive implemented was built of data blocks.

The I/O Latency time along with other statistics mentioned above was (taken from the source):

@@

Over the course of Time the latency of RAM has improved, but compared to the bandwidth and capacity it has
been on the very low side. According to a research done by Kevin Chang from CMU, in his paper 'Understanding
and Improving the Latency of D-RAM based Memory Systems', in the period of 1999 - 2017, the capacity of
RAM has increased a gigantic 128 times, where as the Bandwidth of a RAM has increased a good enough 20 times,
however the improvenement in latency haa only a factor of 1.3x. We can safely assume that the latency of a RAM
has been consistent compared to the latency of Hard Drive over time or the CPU.

## ##

@@

As of 2012 these were some of the Latencies, that includes different operation on a low level
and at the high level.(Source: }

@@

Explain any discrepancies found here.

\section{Conclusions}
\label{sec:conclusions}

The goal of the project was to find out about the varying latencies of some of the major components of the CPU, with
or without the simulator. By implementing a simulator good enough to differentiate and manage data we were able to
find out more about the

>>>>

Seek time; which is the time it takes for the Read-Write to move up and down the platter.
The Seek time itself is comprised of a series of phase times. The first being the acceleration phase,
the read write arm starts to move, then comes the coasting phase when the arm is moving at full speed. This
is followed by the deceleration phase where the read write arm slows down. Then finally comes the
settling phase as the read write head stops right at the data it needs.
The average seek time of hardrives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
(source: \cite{tbree} )

The last part of total I/O latency for hard drive is the transfer time, which is the time it takes the
data to be transfer from or to the Read-Write head. Compared to Rotational Latency and seek time, transfer
time is insignificant, given that the mechanical movements take more time than data movement which is more of
movement of energy packets.

@@

In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the \cite{tbree},
two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
data being dealt with was 100 MB. However, another controller in the comparison was whether the data reading or writing
was sequential or random. Sequential meant that large chunks of data are just being read and written which are in
sequence, whereas the Random transfer just meant Random data was selected from a read issue some where on the disk,
which is how the data for applications work. Mind it that the data we represented in the simulator was also sequential
and not random, as the hardrive implemented was built of data blocks.

The I/O Latency time along with other statistics mentioned above was (taken from the source \cite{tbree}):

@@

Over the course of Time the latency of RAM has improved, but compared to the bandwidth and capacity it has
been on the very low side. According to a research done by Kevin Chang from CMU, in his paper 'Understanding
and Improving the Latency of D-RAM based Memory Systems'(\cite {DRAMTimings}) in the period of 1999 - 2017,
the capacity of RAM has increased a gigantic 128 times, where as the Bandwidth of a RAM has increased a
good enough 20 times, however the improvenement in latency had  only a factor of 1.3x. We can safely assume
that the latency of a RAM has been consistent compared to the latency of Hard Drive over time or the CPU.

@@

As of 2012 these were some of the Latencies, that includes different operation on a low level
and at the high level.(Source: \cite {Latencies})

@@

The goal of the project was to find out about the varying latencies of some of the major components of the CPU, with
or without the simulator. By implementing a simulator good enough to differentiate and manage data we were able to
find out more about the memory management in an environment ran by threads. However, since there exists some complex
and deep details and components in the real world units that we represented, we were only able to implement as much
as the basic level of each component.
The real world data that we represented made up most of our information in the paper, but that was along with the
fact that the data we imported, was a combination of several different projects and researchs.
Starting from the hard drive we did not represent the sectors, platters and other mechanical components, instead
we implemented the HDD using data blocks, thus we were only able to gather stats for a sequential read and write.
A similar implementation of RAM meant that we did not explicitly render the memory cells in the RAM but only the
data blocks which act as those cells.
For the CPU we only represented a single level of cache a L1, while almost all the modern CPUS work on dual level
of cache.
Being programmed in java our simulator was bounded by the high level barrier, so to represent the low
level of memory management, we did have to implement delays ourselves. Our delays were then rationalized according
to the time it takes in the low level. 
A lot that was not gathered through the simulator, a deep analysis of sources did cover for that. Especially getting
to know the latencies of new combinations of components that we did not implement in out simulator helped us to a
more realistic approach. The research itself was based upon a latency simulator with a concentration on what makes
up the delays in a real world programming system.

\section{Conclusions}

\label{sec:conclusions}

Memory Management within a memory system is a complex set of organizations of several different peices, which
themselves are a set of several different elements. So if we are to study the time delays as a superset of
memory latencies, we cannot just discuss the general latencies of each component seperately. But one must
realize that the components are very closely linked with each other. An example of this would be that if a
a CPU references either  cache and it hits, that would be an ideal situation for the a real model, but if does
not find what it needs then it has to ref the memory or eventually the Disks, resulting in so many differnt
factors for delays, including but not only, the CAS latency of the RAM, the HDD I/O latency, which further
includes the delays of rotation, read write head, settling time etc.
